---
title: "20221117_HIVHEPC_exploration"
output: html_document
date: "2022-11-17"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

```{r working directory}
# if necessary,
# remove.packages("ergm")
# remotes::install_version("ergm", version = "3.11.0", repos = "http://cran.us.r-project.org")

# libraries
library(here)
library(ergm)  # NOTE: can't run on ergm 4.0 or higher because of deprecated functons
library(sna)
library(network)
library(rle)

# loading data
node_data = read.csv(file = here("Data", "HIVHEPC_node_data.csv"))
tie_data = read.csv(file = here("Data", "HIVHEPC_tie_data.csv"))
```

## Node data inspection

### Descriptives

```{r descriptives}
# summary of the node data
summary(node_data)

# some frequencies
plot(table(node_data$Sector), type = "h", main = "Frequencies of reported Sector")
table(node_data$Sector)

plot(table(node_data$ActualSector), type = "h", main = "Frequencies of actual Sector")
table(node_data$ActualSector)

plot(table(node_data$OrgType), type = "h", main = "Frequencies of organisation type (OrgType)")
table(node_data$OrgType)

plot(table(node_data$Type), type = "h", main = "Frequencies of organisation type (Type)")
table(node_data$Type)

plot(table(node_data$OrgSize), type = "h", main = "Frequencies of organisation size")
table(node_data$OrgSize)

plot(table(node_data$Level), type = "h", main = "Frequencies of organisation level")
table(node_data$Level)

plot(table(node_data$FundStatus), type = "h", main = "Frequencies of funding status")
table(node_data$FundStatus)

plot(table(node_data$Resp), type = "h", main = "Frequencies of respondents")
table(node_data$Resp)
```

Off brief inspection, it seems like there seems to be more organisations who actually research HIV than what is reported.

Looking at the proportions of organisation type, Peer organisations and NGOs are the most common (CBO and NGO-Health second most common) and Govt organisations are the least common (Hospitals and CALD-ATSI are the second least common).

One thing to note is the 'Hierarchy' as listed in the excel datasheet.

```{r hierarchy table}
# table here because rmd is not very table friendly
hierarchy = data.frame(Number = 1:3, 
                       Type = c("Research,Govt", "CBO,PeerOrgs,CALD-ATSI", "Health,Hospitals,NGO,NGO-Health"))

table(node_data$Type)

knitr::kable(hierarchy)
```


One more thing to note is that **Org30, Org37, Org42, and Org47** are all organisations that are not on the name interpreter list (i.e., they would have 0 indegree). The following deletes these from both data objects.

```{r delete boundary misspecification nodes}
nodes.df = node_data
ties.df = tie_data

renegade4 <- c("Org30", "Org37", "Org42", "Org47")
nodes.df <- nodes.df[-which(nodes.df[,1] %in% renegade4),]

# there should be no rows in which the renegade 4 receive nominations
length(which(ties.df[,2] %in% renegade4))==0

ties.df <- ties.df[-which(ties.df[,1] %in% renegade4),]

nodes.df$hierarchy <- 0
for(i in c(1:nrow(hierarchy))){
nodes.df[which(nodes.df$Type %in% unlist(sapply(hierarchy$Type, strsplit, ",")[i],
                                         use.names = FALSE)),
         which(colnames(nodes.df)=="hierarchy")] <- i
}

node_data = nodes.df
tie_data = ties.df

head(nodes.df)
```



### Closer look into organisation indicators

```{r inspecc orgs}
# let's look closer into the indicators for the organisation type
orgInds = node_data[, 7:14]

# can organisations only be one type?
sum(rowSums(orgInds) == 1)

# Answer: No.

# let's check the ones that have both types
orgInds[rowSums(orgInds) != 1,]

# they're all CALD-ATSI.
```

So this inspection was to check if there was mutual exclusivity in some of these organisation types (e.g., A hospital probably can't be a community-based organisation (CBO) at the same time).

It seems that the only exception is when they are CALD-ATSI.

**Q:** Would it be better to keep the organisation type as mutually exclusive and have a separate indicator for whether they accommodate CALD-ATSI services?

## Tie data inspection

```{r renaming variables}
# let's first rename the variables so it's clearer what they represent.
# Taken from the survey,
tieTypeCodebook = data.frame(varName = colnames(tie_data[,-c(1,2)]),
                             tieMeaning = c("3yr_Collaborate",
                                            "3yr_Shared_insight",
                                            "12mo_Insight_send",
                                            "12mo_Insight_receive",
                                            "12mo_Org_planning_send",
                                            "12mo_Org_planning_receive",
                                            "12mo_Acted_on_received_info",
                                            "12mo_Shared_resources_with",
                                            "12mo_Collaborate_advocacy"))

# print that table
knitr::kable(tieTypeCodebook)

# for transparency's sake let's rename these variables
colnames(tie_data)[-c(1,2)] = tieTypeCodebook$tieMeaning
```

```{r tie data}
## for ease's sake let's start by binarising the tie
##choose the cutoff (tie weights of this number and above are considered a tie)
netlvl <- 2

tie_data_bin = tie_data
tie_ind = tie_data_bin[,-c(1,2)] > (netlvl - 1)
tie_data_bin[, -c(1,2)] = as.numeric(tie_ind)

# a check to make sure they're identical (i.e., the 0s are exactly right), this should be 0.
sum((tie_data == 0) != (tie_data_bin == 0))

# let's start with an example with Q1.
# make an empty data structure
adjMat = matrix(0, nrow = nrow(node_data), ncol = nrow(node_data))
rownames(adjMat) = node_data$ID
colnames(adjMat) = rownames(adjMat)


## Adjacency array
adjArray = array(data = 0, 
                 dim = c(nrow(node_data), nrow(node_data), nrow(tieTypeCodebook)),
                 dimnames = list(c(node_data$ID),
                                 c(node_data$ID),
                                 c(tieTypeCodebook$tieMeaning)))


# loop to translate edge list to adjarray

# loop over all tie types
for( tieType in 1:nrow(tieTypeCodebook) ){

  # chosen tie type index
  tieInds = tie_data_bin[, 2 + tieType]    # the 2+ is to account for the FROM and TO columns
  
  # and loop over all entries
  for( edgeInd in 1:nrow(tie_data_bin) ){
  
    # grab the sender and receiver
    sender = tie_data_bin$FROM[edgeInd]
    receiver = tie_data_bin$TO[edgeInd]
  
    # start filling in the matrix
    adjArray[sender,receiver, tieType] = tieInds[edgeInd]
  }
}
  

# check to make sure the number of edges match
dataEdges = as.numeric(apply(tie_data_bin[,-c(1,2)], MARGIN = 2, FUN = sum))
arrayEdges = as.numeric(apply(adjArray, MARGIN = c(3), FUN = sum))

# any mismatch? should be 0
sum(dataEdges != arrayEdges)

```

### Create new covariates
```{r create new covariates}
library(dplyr)
library(forcats)
head(node_data)

node_data <- node_data %>%
  mutate(funder = fct_collapse(FundStatus,
                               Recipient = c("Funded", "Unfunded"),
                               Funder = "Funder")) %>%
  # mutate(HCV = fct_collapse(ActualSector,
  #                              No = "HIV",
  #                              Yes = c("Both", "HCV"))) %>%
  mutate(HIV = fct_collapse(ActualSector,
                               No = "HCV",
                               Yes = c("Both", "HIV"))) %>%
  mutate(HIVonly = fct_collapse(ActualSector,
                               Yes = "HIV",
                               No = c("Both", "HCV"))) %>%
  mutate(Both = fct_collapse(ActualSector,
                               Yes = "Both",
                               No = c("HCV", "HIV"))) %>%
  mutate(HCV = fct_collapse(ActualSector,
                               No = "HIV",
                               Yes = c("Both", "HCV"))) %>%
  mutate(HCVonly = fct_collapse(ActualSector,
                               Yes = "HCV",
                               No = c("Both", "HIV")))
```


### Covariate formatting

```{r covariates}
## Some monadic attributes
# turn into factors to make sure levels are consistent and then to numeric because ergms don't like 'factors'
# Note: factor indexing starts at 1. This is only annoying when we're dealing with BINARY data. Categorical data should be fine.
sector = as.factor(node_data$Sector)
actualSector = as.factor(node_data$ActualSector)
orgType = as.factor(node_data$OrgType)
orgSize = as.factor(node_data$OrgSize)
type = as.factor(node_data$Type)
level = as.factor(node_data$Level)
fundStatus = as.factor(node_data$FundStatus)
funder = as.factor(node_data$funder)
HCV = as.factor(node_data$HCV)
HIV = as.factor(node_data$HIV)
HCVonly = as.factor(node_data$HCVonly)
HIVonly = as.factor(node_data$HIVonly)
Both = as.factor(node_data$Both)
hierarchy = as.factor(node_data$hierarchy)
CALD = as.numeric(node_data$CALD.ATSI)

# check levels
levels(sector)
levels(actualSector)
levels(orgType)
levels(orgSize)
levels(type)
levels(level)
levels(fundStatus)
levels(funder)
levels(HCV)
levels(HCVonly)
levels(HIV)
levels(HIVonly)
levels(Both)
levels(hierarchy)


# save as numeric
sector = as.numeric(sector)
actualSector = as.numeric(actualSector)
orgType = as.numeric(orgType)
orgSize = as.numeric(orgSize)
type = as.numeric(type)
level = as.numeric(level)
national = abs(as.numeric(level)-2)
fundStatus = as.numeric(fundStatus)
funder = as.numeric(funder)-1
HIV = abs(as.numeric(HIV)-2)
HIVonly = abs(as.numeric(HIVonly)-1)
HCV = abs(as.numeric(HCV)-2)
HCVonly = abs(as.numeric(HCVonly)-1)
Both = abs(as.numeric(Both)-2)
hierarchy = 4 - as.numeric(hierarchy)

## edge covariate for 'high trust organisational ties' (i.e., 2s in the edge list)
## Adjacency array
highTies = array(data = 0, 
                 dim = c(nrow(node_data), nrow(node_data), nrow(tieTypeCodebook)),
                 dimnames = list(c(node_data$ID),
                                 c(node_data$ID),
                                 c(tieTypeCodebook$tieMeaning)))


# loop to translate edge list to adjarray

# loop over all tie types
for( tieType in 1:nrow(tieTypeCodebook) ){

  # chosen tie type index
  tieInds = tie_data[, 2 + tieType]    # the 2+ is to account for the FROM and TO columns
  
  # and loop over all entries
  for( edgeInd in 1:nrow(tie_data_bin) ){
  
    # grab the sender and receiver
    sender = tie_data_bin$FROM[edgeInd]
    receiver = tie_data_bin$TO[edgeInd]
  
    # start filling in the matrix
    highTies[sender,receiver, tieType] = tieInds[edgeInd]
  }
}

# recode all 0s and 1s in highTies to be 0s and 2s to be 1s
highTies[highTies <= 1] = 0
highTies[highTies == 2] = 1

# check to make sure the number of edges match
highTieData = tie_data[, -c(1,2)] == 2

dataHighs = as.numeric(apply(highTieData, MARGIN = 2, FUN = sum))
arrayHighs = as.numeric(apply(highTies, MARGIN = c(3), FUN = sum))

# any mismatch? should be 0
sum(dataHighs != arrayHighs)

# highTies can be used as an edge covariate by taking out the array with the selected tie type (e.g., 1 is collaborate high ties)

```



### Network inspection

```{r network check, fig.height = 6, fig.width = 12}

# loop to check all networks
par(mfrow = c(1,3))

for(tieType in 1:9){
  
  # choose adjmat
  adjMat = adjArray[,,tieType]
  
  # plot the network
  gplot(adjMat, 
        gmode = "digraph",
        edge.col = "grey",
        usecurve = TRUE,
        edge.curve = 0.05,
        arrowhead.cex = 0.5,
        vertex.cex = sna::degree(adjMat) * 0.075,
        main = tieTypeCodebook$tieMeaning[tieType])

  # bit difficult to see the plot, so maybe let's try a degree distribution
  plot(table(rowSums(adjMat)), type = "h", main = "Outdegree distribution", xlab = "Outdegree", ylab = "Frequency")
  lines(x = as.numeric(names(table(rowSums(adjMat)))), y = as.numeric(table(rowSums(adjMat))), col = "red") 

  plot(table(colSums(adjMat)), type = "h", main = "Indegree distribution", xlab = "Indegree", ylab = "Frequency")
  lines(x = as.numeric(names(table(colSums(adjMat)))), y = as.numeric(table(colSums(adjMat))), col = "red") 


  # some censes
  dyad.census(adjMat)
  triad.census(adjMat)

}

```

```{r choose tie}
  
### TODO: Choose one particular adjacency matrix
chosenTieType = 5

# 1 = 3yr_Collaborate
# 2 = 3yr_Shared_insight
# 3 = 12mo_Insight_send
# 4 = 12mo_Insight_receive
# 5 = 12mo_Org_planning_send
# 6 = 12mo_Org_planning_receive
# 7 = 12mo_Acted_on_received_info
# 8 = 12mo_Shared_resources_with
# 9 = 12mo_Collaborate
## These can also be used as edge covariates

# a check
paste("Your chosen tie type is", tieTypeCodebook$tieMeaning[chosenTieType])

# and we take from the adj array
adjMat = adjArray[,,tieTypeCodebook$tieMeaning[chosenTieType]]
  
```


Just off a brief inspection, it seems that the indegree distribution is not nearly as skewed as the outdegree distribution. There are only a few organisations with very high outdegree.

## ERGM time, Missing.

### What parts are missing?

On thinking, due to being a name interpreter (i.e., list of 44 shown organisations), some null ties can be confirmed to be true null ties (i.e., the organisations have been given the option to nominate the organisation and chose not to).

However, not all organisations responded to the survey, so technically speaking the outdegrees of the non-respondents should all be considered missing.

Some tie types may facilitate a reciprocal relationship in terms of e.g., 'they shared their insight with me, thus they would acknowledge that they did indeed share their insight with the focal organisation'.

One other consideration is that 4 organisations (Org30, Org37, Org42, Org47) were not in the name interpreter, so they could not have received incoming ties.

```{r check miss, eval=FALSE}
## non-respondents have 0 outdegree
# this is indeed the case, should be 0 for no discrepancy
sum((node_data$Resp=="N") != as.logical(rowSums(adjMat) == 0))

# another way to check
setequal(node_data$Resp=="N", as.logical(rowSums(adjMat) == 0))

## organisations not in the name interpreter have 0 indegree
print(node_data$ID[as.logical(colSums(adjMat) == 0)])
```

### Set missing values

```{r make missing}
## Adding missing tie variables to the adjacency matrix as described above
missAdj = adjMat

# Set row of all nonrespondents to missing
nonResp = node_data$ID[node_data$Resp=="N"]
missAdj[nonResp,] = NA

# organisations not in the name interpreter
# noIntOrg = c("Org30", "Org37", "Org42", "Org47")

# missAdj[, noIntOrg] = NA

# check how many missing tie variables there are
table(missAdj, useNA = "always")

```


### Miss ERGM Bayes has entered the building
```{r missERGMBayes, warning = FALSE}
# load routine
source(here("Code", "miss_ERGM_Bayes.r"))

# change the outcome object to use the matrix with missingness
net = as.network(missAdj)


```


```{r moremissbayes}

modelnum <- 106
modelname <- paste0("model_",modelnum)

source(here("Code", "Models", paste0(modelname,".R")))

startTime = Sys.time()
#more Miss Bayes

#burn in 
missModelbrn = miss_ERGM_Bayes(formula = currentmodel,
                                tot.iterations = 2000,
                                tuning.const = 0.5,
                                thinning = 1,
                                burnin = 1,
                                save.nets = FALSE,
                                num.nets = 20,
                                printFreq = 500,
                                MultFac.aux = 200,
                                MultFac.miss = 200,
                                diagonal.proposal = FALSE,
                                warm.burnin = 5000,
                                warm.thin = 5000)



# save covariance matrix for proposal covariance
prop.var <- cov( missModelbrn$Post.theta[1:2000,] ) 

# save the end point of the burn in model
start.theta <- missModelbrn$Post.theta[2000,]


# actual model
missModelMain = miss_ERGM_Bayes(formula = currentmodel,
                             tot.iterations = 3000,
                             tuning.const = 0.5,
                             thinning = 1,
                             burnin = 1,
                             save.nets = TRUE,
                             num.nets = 100,
                             printFreq = 500,
                             MultFac.aux = 200,
                             MultFac.miss = 200,
                             diagonal.proposal = FALSE,
                             warm.burnin = 1000,
                             warm.thin = 1000,
                             sigma.epsilon = prop.var,
                             theta = start.theta)

endTime = Sys.time()

print(endTime - startTime)

# check model results
plot(ts(missModelMain$Post.theta[,c(1:5)]))
plot(ts(missModelMain$Post.theta[,c(6:10)]))
plot(ts(missModelMain$Post.theta[,c(11:15)]))
plot(ts(missModelMain$Post.theta[,c(16:20)]))
plot(ts(missModelMain$Post.theta[,c(20:23)]))

#table
res.table <- creatTab(missModelMain$Post.theta)
res.table

# write.csv(missModelMain, file='Markov_alt_shareresoruces.csv')

saveRDS(missModelMain,file = paste0("results2_net_",chosenTieType, "_lvl_",netlvl,"_", modelname,"_",format(Sys.time(), "%d_%m_%y_%H%M"),".RDS"))

```

### Make a nice table for the results
```{r tablemaker}
library(gt)

sigmarker <- which((res.table[,4]/res.table[,5]) > 0)



res.table.1 <- res.table[c(1:ceiling(nrow(res.table)/2)),]
res.table.2 <- res.table[c(ceiling(nrow(res.table)/2) + 1):nrow(res.table),]
res.table.1 <- res.table.1 %>%
  gt()
res.table.2 <- res.table.2 %>%
  gt()

listed_tables <- list(res.table.1, res.table.2)

gt_two_column_layout(listed_tables)

res.table %>%
  gt() #%>%
  # gt_highlight_rows(
  #    rows = 5,
  #    fill = "lightgrey",
  #    bold_target_only = TRUE
  #  )

```
