---
title: "GOF for missBayes"
author: "Jonathan J"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###Load data
```{r loaddata}
# missModelMain
```
## Goodness-of-fit: comparing imputed and predicted
We can compare the imputed networks `object$mix.nets` with the networks `aux.nets` predicted by the model manually using desnity, dyad, or triad census. Any number of functions of the imputed and predicted networks can be calculated, to investigate how the model fits a population network (Stys et al. 2020)

###GOF for degree distribution
The GOF distributions for degree are given by
```{r gofdegs}
# library(ergm)
gofdegs <- gof.degree(missModelMain)
```

A plot function superimposes the imputed and predicted
```{r gofdegs2}

plot.gof.degree(gofdegs,max.deg.plot.ind = 11)
```

###Checking imputed degree distributions against the observed degree distributions for respondents
If data are missing at random, we would expect that the outdegree distribution for the imputed networks is similar to the outdegree distributions for the respondents.

```{r gofdegs3}

# is.resopndent <- sapply((node_data$Resp=="Y"), as.numeric)
is.resopndent <- (node_data$Resp=="Y")

tempdeg <- gof.obs.degree(obs.net=net,mat=missModelMain,is.resopndent=is.resopndent ,directed=TRUE)
par(mfrow=c(1,2))
boxplot(tempdeg$mix.indeg[,1:10],ylim=range(tempdeg$mix.indeg[,1:10],tempdeg$aux.indeg[1:10] ), main='indegree' ,names=c(0:9))
lines(tempdeg$aux.indeg[1:10],col='red')
boxplot(tempdeg$mix.oudeg[,1:10],ylim=range(tempdeg$mix.oudeg[,1:10],tempdeg$aux.oudegg[1:10] ) ,main='outdegree',names=c(0:9))
lines(tempdeg$aux.oudeg[1:10],col='red')
lines(table(rowSums(adjMat)[is.resopndent]+1)/sum(is.resopndent))
```


##GOF for geodesic distances
The GOF distributions for distances are given by
```{r gofgoedist}
gofdist <- gof.dist(missModelMain, directed=TRUE , max.path = 35)
```

A plot function superimposes the imputed and precicted
```{r gofgoedist2}
plot.gof.dist(gofdist,max.dist.plot.ind = 35,do.inf.dist=FALSE)
```

##GOF for edge-wise shared partners
The GOF distributions for edge-wise shared partners are given by
```{r gof.gwesp}
gofesp <- gof.esp(ans=missModelMain,directed=TRUE)
```
A plot function superimposes the imputed and precicted
```{r gof.gwesp.2}
plot.gof.esp(gofesp= gofesp)
```

##Manually inspect the triad census
Calculate the triad census for the predicted networks and the imputed networks
```{r triadcensus}
TriadPred <- triad.census(missModelMain$aux.nets)
TriadImputed <- triad.census(missModelMain$mix.nets)
```
Compare, side by side, the triad counts of the predicted networks and the imputed networks.
```{r triadcensus2}
par( mfrow = c(4,4), oma = c(0,4,0,0) + 0.1,
      mar = c(5,0,1,1) + 0.1)# oma sets outer margins and mar the margins
for (j in c(1:16)){
boxplot(TriadPred[,j],TriadImputed[,j],names = c('Pred','imp'), xlab=colnames(TriadPred)[j], yaxt='n' )
 abline(h=mean(TriadImputed[,j]),col='red',lty=3) 
}
```